Plant Trait Imputation

### 2024-06-03 code/L1/L1_LastStep_DiversityInputObjects.Rmd
update function to assign cellid as row names

```{r}
# Define the function
create_presence_absence_matrix <- function(resolution_meters, species_sf) {
  # Make Grid
  TAGrid <- TApoly %>%
    st_make_grid(cellsize = c(resolution_meters)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number(),
           centroid = st_centroid(geometry)) %>%
    mutate(longitude = st_coordinates(centroid)[, 1],
           latitude = st_coordinates(centroid)[, 2]) %>%
    dplyr::select(-centroid)

  # Join with species data and filter out NA species
  species_grid <- TAGrid %>%
    st_join(species_sf) %>%
    filter(!is.na(species)) %>%
    dplyr::select(cellid, longitude, latitude, species)

  # Create Presence-Absence Matrix
  presence_absence_matrix <- species_grid %>%
    st_set_geometry(NULL) %>%
    mutate(presence = 1) %>%
    pivot_wider(names_from = species, values_from = presence, values_fill = list(presence = 0)) %>%
    arrange(cellid) %>%
    dplyr::select(cellid, longitude, latitude, everything())

  # Convert to matrix and set row names
  presence_absence_matrix <- as.data.frame(presence_absence_matrix)
  rownames(presence_absence_matrix) <- presence_absence_matrix$cellid
  presence_absence_matrix <- presence_absence_matrix %>%
    dplyr::select(-cellid) %>%
    as.matrix()

  return(presence_absence_matrix)
}
```

@@

### Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final_100km <- plant_PAM_100km[, colSums(plant_PAM_100km) != 0]
PAM_frugivore_site_final_100km <- frugivore_PAM_100km[, colSums(frugivore_PAM_100km) != 0]

#Save coordinates for later
site_loc_key_plant_100km <- PAM_plant_site_final_100km[,1:2]
site_loc_key_frugivore_100km <- PAM_frugivore_site_final_100km[,1:2]

@@

colnames_plant_100km <- colnames(PAM_plant_site_final_100km)
colnames_frugivore_100km <- colnames(PAM_frugivore_site_final_100km)
```

>>>>

```{r}
create_presence_absence_matrix <- function(resolution_meters, species_sf) {
  # Make Grid
  TAGrid <- TApoly %>%
    st_make_grid(cellsize = c(resolution_meters)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())

  # Join with species data and filter out NA species
  species_grid <- TAGrid %>%
    st_join(species_sf) %>%
    filter(!is.na(species)) %>%
    dplyr::select(cellid, species, geometry)

  # Extract the coordinates for each grid cell
  species_grid_coords <- species_grid %>%
    st_centroid() %>%
    st_coordinates() %>%
    as.data.frame() %>%
    rename(Longitude = X, Latitude = Y)

  # Combine coordinates with the species grid
  species_grid <- bind_cols(species_grid, species_grid_coords)

  # Create Presence-Absence Matrix
  presence_absence_matrix <- species_grid %>%
    st_set_geometry(NULL) %>%
    mutate(presence = 1) %>%
    pivot_wider(names_from = species, values_from = presence, values_fill = list(presence = 0))

  # Convert to matrix
  presence_absence_matrix_matrix <- as.matrix(presence_absence_matrix)

  # Set row names as "cell_rowNumber"
  rownames(presence_absence_matrix_matrix) <- paste0("cell_", seq_len(nrow(presence_absence_matrix_matrix)))

  # Ensure latitude and longitude are the first columns
  presence_absence_matrix_matrix <- presence_absence_matrix_matrix[, c("Latitude", "Longitude", setdiff(colnames(presence_absence_matrix_matrix), c("Latitude", "Longitude", "cellid")))]

  return(presence_absence_matrix_matrix)
}
```

@@

### Remove the species from PAM that have no occurrences
```{r}
dim(plant_PAM_100km)
dim(frugivore_PAM_100km)
# Remove columns with sum equal to zero
PAM_plant_site_final_100km <- plant_PAM_100km[, colSums(plant_PAM_100km) != 0]
PAM_frugivore_site_final_100km <- frugivore_PAM_100km[, colSums(frugivore_PAM_100km) != 0]

dim(PAM_plant_site_final_100km)
dim(PAM_frugivore_site_final_100km)

# Save coordinates for later
site_loc_key_plant_100km <- PAM_plant_site_final_100km[,1:2]
site_loc_key_frugivore_100km <- PAM_frugivore_site_final_100km[,1:2]

@@

colnames_plant_100km <- colnames(PAM_plant_site_final_100km)
colnames_frugivore_100km <- colnames(PAM_frugivore_site_final_100km)

```
